---
title: "QTM final project"
author: "Naibo(Ray) Hu"
date: "4/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Docs/Babson/QTM3605/Final Project')
```

#Load File
```{r}
year13= read.csv("May13 public.csv")
year15= read.csv("May15 public.csv")
year10= read.csv("Sept10 public.csv")
year17= read.csv("Typology 17 public.csv")

```

# Data Preprocessing
```{r}
library(dplyr)
#2013
anyNA(year13$q36a)
year13_v1= filter(year13, q36a !=9)
year13_v1$lgbtq_marry<- ifelse(year13_v1$q36a== 1 
                               | year13_v1$q36a== 2 ,1,0)
#2015
anyNA(year15$q40)
year15_v1 = filter(year15, q40 != 9)
#1 support 0 oppose
year15_v1$lgbtq_marry<- ifelse(year15_v1$q40== 1 
                            | year15_v1$q40== 2 ,1,0)
#2010
anyNA(year10$q40)
year10_v1= filter(year10, q40 !=9)
year10_v1$lgbtq_marry<- ifelse(year10_v1$q40== 1 
                               | year10_v1$q40== 2 ,1,0)
#2017
anyNA(year17$qa128)
sum(is.na(year17$qa128))

year17_v1= filter(year17, !is.na (year17$qa128))
year17_v1= filter(year17, qa128 !=9)
year17_v1$lgbtq_marry<- ifelse(year17_v1$qa128== 1 
                               | year17_v1$qa128== 2 ,1,0)
```
```{r}
##2013
anyNA(year13_v1$receduc)
anyNA(year13_v1$q58) #religion conflict
anyNA(year13_v1$party)
anyNA(year13_v1$q61) #interaction
sum(is.na(year13_v1$q61)) 
year13_v2 = filter(year13_v1, !is.na(q61), receduc!=9, q58!=9, party!=9, q61!=9)

#2013 interaction (q61)
unique(year13_v2$q61)
year13_v2$interaction <- ifelse(year13_v2$q61 == 1, 1,0) #1-yes 0-no

#2013 religion conflict (q58)
unique(year13_v2$q58)
year13_v2$relig_conflict <- ifelse(year13_v2$q58 <= 2, 1,0) #1-conflict, 0-no conflict

#2013 education 
unique(year13_v2$receduc)
year13_v2$HS_grad <- ifelse(year13_v2$receduc == 1, 1,0) #HS grad or less 
year13_v2$some_coll <- ifelse(year13_v2$receduc == 2, 1,0) #Some college/Associate Degree
year13_v2$coll <- ifelse(year13_v2$receduc ==3, 1,0) #Bachelor's degree or higher

#2013 partisanship
unique(year13_v2$party)
year13_v2$republic <- ifelse(year13_v2$party == 1, 1,0) #Republican
year13_v2$democrat <- ifelse(year13_v2$party == 2, 1,0) #Democrat
year13_v2$other_party <- ifelse(year13_v2$party >= 3, 1,0) #Other: Independent, No preference, Other party 

##2015
anyNA(year15_v1$educ2)
anyNA(year15_v1$q49) #religion conflict
anyNA(year15_v1$party)
anyNA(year15_v1$q52) #interaction
sum(is.na(year15_v1$q52))
year15_v2 = filter(year15_v1, !is.na(q52), educ2!=9, q49!=9, party!=9, q52!=9)

#2015 interaction (q52)
unique(year15_v2$q52)
year15_v2$interaction <- ifelse(year15_v2$q52 == 1, 1,0) #1-yes 0-no

#2015 religion conflict
unique(year15_v2$q49)
year15_v2$relig_conflict <- ifelse(year15_v2$q49 <= 2, 1,0) #1-conflict, 0-no conflict

#2015 education
unique(year15_v2$educ2)
year15_v2$HS_grad <- ifelse(year15_v2$educ2 <= 3, 1,0) #HS grad or less 
year15_v2$some_coll <- ifelse(year15_v2$educ2 == 4 
                              | year15_v2$educ2== 5, 1,0) #Some college/Associate Degree
year15_v2$coll <- ifelse(year15_v2$educ2 >= 6, 1,0) #Bachelor's degree or higher

#2015 partisanship
unique(year15_v2$party)
year15_v2$republic <- ifelse(year15_v2$party == 1, 1,0) #Republican
year15_v2$democrat <- ifelse(year15_v2$party == 2, 1,0) #Democrat
year15_v2$other_party <- ifelse(year15_v2$party >= 3, 1,0) #Other: Independent, No preference, Other party 


```

#National Oppose Percentage
```{r}
# 2010 oppose percentage
count_10=0
for (i in year10_v1$lgbtq_marry){
  if(i==0)
    count_10 = count_10+1
}
print(count_10)
p_10= print(count_10/nrow(year10_v1)*100)

# 2013 oppose percentage
count_13=0
for (i in year13_v1$lgbtq_marry ){
  if(i==0)
    count_13 = count_13+1
}
print(count_13)
p_13=print(count_13/nrow(year13_v1)*100)

#2015 oppose percentage
count_15=0
for (i in year15_v1$lgbtq_marry){
  if(i==0)
    count_15= count_15+1
}
print(count_15)
p_15=print(count_15/nrow(year15_v1)*100)

#2017 oppose percentage
count_17=0
for (i in year17_v1$lgbtq_marry){
  if(i==0)
    count_17=count_17+1 
}
print(count_17)
p_17=print(count_17/nrow(year17_v1)*100)
```

# Line Chart 1 
```{r}
#time series line chart 

# Create the data for the chart.
v <- c(p_10,p_13,p_15,p_17)

# Plot the bar chart.
plot(v,type = "o", col = "blue", xlab = "year", ylab = "Percentage",
     main = "Percentage of people Oppose LGBTQ marriage in U.S.",xaxt = "n",ylim=c(20,60))
axis(1,at=1:4,labels=c("2010","2013","2015","2017"))

```
#states pass legalization locally prior 2015
```{r}
#California 2013 June
ca_10=filter(year10_v1, state==6)
ca_13=filter(year13_v1, state==6)
ca_15=filter(year15_v1, state==6)
ca_17=filter(year17_v1, state==6)

#New York 2011 June
ny_10=filter(year10_v1, state==36)
ny_13=filter(year13_v1, state==36)
ny_15=filter(year15_v1, state==36)
ny_17=filter(year17_v1, state==36)

#Washington 2012 Feb
wa_10=filter(year10_v1, state==53)
wa_13=filter(year13_v1, state==53)
wa_15=filter(year15_v1, state==53)
wa_17=filter(year17_v1, state==53)

#Virginia 2014 Oct
va_10=filter(year10_v1, state==51)
va_13=filter(year13_v1, state==51)
va_15=filter(year15_v1, state==51)
va_17=filter(year17_v1, state==51)

#merge data
pass_10= rbind(ca_10,ny_10, va_10, wa_10)
pass_13= rbind(ca_13,ny_13, va_13, wa_13)
pass_15= rbind(ca_15,ny_15, va_15, wa_15)
pass_17= rbind(ca_17,ny_17, va_17, wa_17)

```

# Percentage

```{r}
#count10
count_pass_10=0
for (i in pass_10$lgbtq_marry){
  if(i==0)
    count_pass_10 = count_pass_10+1
}
print(count_pass_10)
pass_p_10= print(count_pass_10/nrow(pass_10)*100)

#count13
count_pass_13=0
for (i in pass_13$lgbtq_marry){
  if(i==0)
    count_pass_13 = count_pass_13+1
}
print(count_pass_13)
pass_p_13= print(count_pass_13/nrow(pass_13)*100)

#count15
count_pass_15=0
for (i in pass_15$lgbtq_marry){
  if(i==0)
    count_pass_15 = count_pass_15+1
}
print(count_pass_15)
pass_p_15= print(count_pass_15/nrow(pass_15)*100)

#count17
count_pass_17=0
for (i in pass_17$lgbtq_marry){
  if(i==0)
    count_pass_17 = count_pass_17+1
}
print(count_pass_17)
pass_p_17= print(count_pass_17/nrow(pass_17)*100)
```

#States did not pass legalization locally prior 2015
```{r}
#Texas
tx_10=filter(year10_v1, state==48)
tx_13=filter(year13_v1, state==48)
tx_15=filter(year15_v1, state==48)
tx_17=filter(year17_v1, state==48)

#Ohio
oh_10=filter(year10_v1, state==39)
oh_13=filter(year13_v1, state==39)
oh_15=filter(year15_v1, state==39)
oh_17=filter(year17_v1, state==39)

#Michigan
mi_10=filter(year10_v1, state==26)
mi_13=filter(year13_v1, state==26)
mi_15=filter(year15_v1, state==26)
mi_17=filter(year17_v1, state==26)

#Georgia
ga_10=filter(year10_v1, state==13)
ga_13=filter(year13_v1, state==13)
ga_15=filter(year15_v1, state==13)
ga_17=filter(year17_v1, state==13)

#merge data
npass_10= rbind(tx_10,oh_10, mi_10, ga_10)
npass_13= rbind(tx_13,oh_13, mi_13, ga_13)
npass_15= rbind(tx_15,oh_15, mi_15, ga_15)
npass_17= rbind(tx_17,oh_17, mi_17, ga_17)
```

#Percentage Oppose
```{r}
#count10
count_npass_10=0
for (i in npass_10$lgbtq_marry){
  if(i==0)
    count_npass_10 = count_npass_10+1
}
print(count_npass_10)
npass_p_10= print(count_npass_10/nrow(npass_10)*100)

#count13
count_npass_13=0
for (i in npass_13$lgbtq_marry){
  if(i==0)
    count_npass_13 = count_npass_13+1
}
print(count_npass_13)
npass_p_13= print(count_npass_13/nrow(npass_13)*100)

#count15
count_npass_15=0
for (i in npass_15$lgbtq_marry){
  if(i==0)
    count_npass_15 = count_npass_15+1
}
print(count_npass_15)
npass_p_15= print(count_npass_15/nrow(npass_15)*100)

#count17
count_npass_17=0
for (i in npass_17$lgbtq_marry){
  if(i==0)
    count_npass_17 = count_npass_17+1
}
print(count_npass_17)
npass_p_17= print(count_npass_17/nrow(npass_17)*100)
```

#Line Chart 2
```{r}
### plot
xdata <- c(1,2,3,4)
y1 <- c(p_10,p_13,p_15,p_17)
y2 <- c(pass_p_10, pass_p_13,pass_p_15, pass_p_17)
y3 <- c(npass_p_10,npass_p_13,npass_p_15,npass_p_17)
plot(xdata, y1, type="o", col="blue", pch="o", lty=1, ylim=c(20,60),xaxt = "n",xlab = "year", ylab = "Percentage",
     main = "Percentage of people oppose LGBTQ marriage in U.S.")
points(xdata, y2, col="red", pch="*")
lines(xdata, y2, col="red",lty=2)
points(xdata, y3, col="orange",pch="+")
lines(xdata, y3, col="orange", lty=3)
axis(1,at=1:4,labels=c("2010","2013","2015","2017"))
```

# Change of Percentage 
```{r}
##change of percentage 
#2010-2013 state not legalize
ch_1= (npass_p_13- npass_p_10)/npass_p_10
#2013-2015 state not legalize
ch_2= (npass_p_15- npass_p_13)/npass_p_13
#2015-2017 state not legalize
ch_3= (npass_p_17- npass_p_15)/npass_p_15

#2010-2013 National
ch_4= (p_13- p_10)/p_10
#2013-2015 National
ch_5= (p_15- p_13)/p_13
#2015-2017 National
ch_6= (p_17- p_15)/p_15

#2010-2013 state legalize
ch_7= (pass_p_13- pass_p_10)/pass_p_10
#2013-2015 state legalize
ch_8= (pass_p_15- pass_p_13)/pass_p_13
#2015-2017 state legalize
ch_9= (pass_p_17- pass_p_15)/pass_p_15

```


# Convert sex and race to factors
```{r}
library(modelsummary)

year13_v2$racecmb <- as.factor(year13_v2$racecmb)
year13_v2$sex <- as.factor(year13_v2$sex)

str(year13_v2)
```

#Interaction
```{r}

model_interaction_13 <- list()
model_interaction_13[['interaction1']] = lm(lgbtq_marry~interaction, data = year13_v2)
model_interaction_13[['interaction2']] = lm(lgbtq_marry~interaction+racecmb+sex+age, data = year13_v2,)
modelsummary(model_interaction_13, stars = TRUE, title = "Regressions - interactions")

```

# Religious Conflict
```{r}
model_relig_conflict_13 <- list()
model_relig_conflict_13[['conflict1']] = lm(lgbtq_marry~relig_conflict, data = year13_v2)
model_relig_conflict_13[['conflict2']] = lm(lgbtq_marry~relig_conflict+racecmb+sex+age, data = year13_v2) #control for demographics
modelsummary(model_relig_conflict_13, stars = TRUE, title = "Regressions - religion conflicts")
```

# Education
```{r}
model_education_13 <- list()
model_education_13[['HS_grad1']] = lm(lgbtq_marry~HS_grad, data = year13_v2)
model_education_13[['HS_grad2']] = lm(lgbtq_marry~HS_grad+racecmb+sex+age, data = year13_v2) #control for demographics
model_education_13[['some_coll1']] = lm(lgbtq_marry~some_coll, data = year13_v2)
model_education_13[['some_coll2']] = lm(lgbtq_marry~some_coll+racecmb+sex+age, data = year13_v2) #control for demographics
model_education_13[['coll1']] = lm(lgbtq_marry~coll, data = year13_v2)
model_education_13[['coll2']] = lm(lgbtq_marry~coll+racecmb+sex+age, data = year13_v2) #control for demographics
modelsummary(model_education_13, stars = TRUE, title = "Regressions - education levels")

```

#Partisanship
```{r}
model_party_13 <- list()
model_party_13[['Republican1']] = lm(lgbtq_marry~republic, data = year13_v2)
model_party_13[['Republican2']] = lm(lgbtq_marry~republic+racecmb+sex+age, data = year13_v2) #control for demographics
model_party_13[['Democrat1']] = lm(lgbtq_marry~democrat, data = year13_v2)
model_party_13[['Democrat2']] = lm(lgbtq_marry~democrat+racecmb+sex+age, data = year13_v2) #control for demographics
model_party_13[['Other1']] = lm(lgbtq_marry~other_party, data = year13_v2)
model_party_13[['Other2']] = lm(lgbtq_marry~other_party+racecmb+sex+age, data = year13_v2) #control for demographics
modelsummary(model_party_13, stars = TRUE, title = "Regressions - partisanships")

```

#Check Missing Values & Convert to factors
```{r}
year15_v2$racecmb <- as.factor(year15_v2$racecmb)
year15_v2$sex <- as.factor(year15_v2$sex)

str(year15_v2)
```

#Interaction
```{r}
model_interaction_15 <- list()
model_interaction_15[['interaction1']] = lm(lgbtq_marry~interaction, data = year15_v2)
model_interaction_15[['interaction2']] = lm(lgbtq_marry~interaction+racecmb+sex+age, data = year15_v2) #control for demographics 
modelsummary(model_interaction_15, stars = TRUE, title = "Regressions - interactions")
```

# Religious Conflict 
```{r}

model_relig_conflict_15 <- list()
model_relig_conflict_15[['conflict1']] = lm(lgbtq_marry~relig_conflict, data = year15_v2)
model_relig_conflict_15[['conflict2']] = lm(lgbtq_marry~relig_conflict+racecmb+sex+age, data = year15_v2)
modelsummary(model_relig_conflict_15, stars = TRUE, title = "Regressions - religion conflicts")

```

# Education
```{r}
model_education_15 <- list()
model_education_15[['HS_grad1']] = lm(lgbtq_marry~HS_grad, data = year15_v2)
model_education_15[['HS_grad2']] = lm(lgbtq_marry~HS_grad+racecmb+sex+age, data = year15_v2) #control for demographics
model_education_15[['some_coll1']] = lm(lgbtq_marry~some_coll, data = year15_v2)
model_education_15[['some_coll2']] = lm(lgbtq_marry~some_coll+racecmb+sex+age, data = year15_v2) #control for demographics
model_education_15[['coll1']] = lm(lgbtq_marry~coll, data = year15_v2)
model_education_15[['coll2']] = lm(lgbtq_marry~coll+racecmb+sex+age, data = year15_v2) #control for demographics
modelsummary(model_education_15, stars = TRUE, title = "Regressions - education levels")
```

# Partisianship
```{r}
model_party_15 <- list()
model_party_15[['Republican1']] = lm(lgbtq_marry~republic, data = year15_v2)
model_party_15[['Republican2']] = lm(lgbtq_marry~republic+racecmb+sex+age, data = year15_v2) #control for demographics
model_party_15[['Democrat1']] = lm(lgbtq_marry~democrat, data = year15_v2)
model_party_15[['Democrat2']] = lm(lgbtq_marry~democrat+racecmb+sex+age, data = year15_v2) #control for demographics
model_party_15[['Other1']] = lm(lgbtq_marry~other_party, data = year15_v2)
model_party_15[['Other2']] = lm(lgbtq_marry~other_party+racecmb+sex+age, data = year15_v2) #control for demographics
modelsummary(model_party_15, stars = TRUE, title = "Regressions - partisanships")
```

